### 2022-10-15 draft.tex
Deploying to gh-pages from @ gottacatchenall/ms_false_negatives@80920980ebd5187ab667cfd01f41e41cfcdd9132 🚀

%!TEX TS-program = xelatex
\documentclass[11pt]{article}

\usepackage[english]{babel}

\usepackage{amsmath,amssymb,amsfonts}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{stix2}
\usepackage[scaled]{helvet}
\usepackage[scaled]{inconsolata}

\usepackage{lastpage}

\usepackage{setspace}

\usepackage{ccicons}

\usepackage[hang,flushmargin]{footmisc}

\usepackage{geometry}

\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}

\usepackage{fancyhdr}
\renewcommand{\headrulewidth}{0pt}\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

\makeatletter
\newcounter{tableno}
\newenvironment{tablenos:no-prefix-table-caption}{
  \caption@ifcompatibility{}{
    \let\oldthetable\thetable
    \let\oldtheHtable\theHtable
    \renewcommand{\thetable}{tableno:\thetableno}
    \renewcommand{\theHtable}{tableno:\thetableno}
    \stepcounter{tableno}
    \captionsetup{labelformat=empty}
  }
}{
  \caption@ifcompatibility{}{
    \captionsetup{labelformat=default}
    \let\thetable\oldthetable
    \let\theHtable\oldtheHtable
    \addtocounter{table}{-1}
  }
}
\makeatother

\usepackage{array}
\newcommand{\PreserveBackslash}[1]{\let\temp=\\#1\let\\=\temp}
\let\PBS=\PreserveBackslash

\usepackage[breaklinks=true]{hyperref}
\hypersetup{colorlinks,%
citecolor=blue,%
filecolor=blue,%
linkcolor=blue,%
urlcolor=blue}
\usepackage{url}

\usepackage{caption}
\setcounter{secnumdepth}{0}
\usepackage{cleveref}

\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth
\else\Gin@nat@width\fi}
\makeatother
\let\Oldincludegraphics\includegraphics
\renewcommand{\includegraphics}[1]{\Oldincludegraphics[width=\maxwidth]{#1}}

\usepackage{longtable}
\usepackage{booktabs}

\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}

\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[3] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1 \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces\fi
  % set entry spacing
  \ifnum #2 > 0
  \setlength{\parskip}{#2\baselineskip}
  \fi
 }%
 {}
\usepackage{calc} % for \widthof, \maxof
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\maxof{\widthof{#1}}{\csllabelwidth}}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth}{#1}}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}\geometry{verbose,letterpaper,tmargin=2.2cm,bmargin=2.2cm,lmargin=2.2cm,rmargin=2.2cm}

\usepackage{lineno}
\usepackage[nolists,noheads]{endfloat}

\pagestyle{plain}

\tolerance=1
\emergencystretch=\maxdimen
\hyphenpenalty=10000
\hbadness=10000

\doublespacing

\fancypagestyle{normal}
{
  \fancyhf{}
  \fancyfoot[R]{\footnotesize\sffamily\thepage\ of \pageref*{LastPage}}
}
\begin{document}
\raggedright
\thispagestyle{empty}
{\Large\bfseries\sffamily The missing link: discerning true from false
negatives when sampling species interaction networks}
\vskip 5em

%
\href{https://orcid.org/0000-0002-6506-6487}{Michael D.\,Catchen}%
%
\,\textsuperscript{1,2}\quad %
\href{https://orcid.org/0000-0002-0735-5184}{Timothée\,Poisot}%
%
\,\textsuperscript{3,2}\quad %
\href{https://orcid.org/0000-0002-6004-4027}{Laura\,Pollock}%
%
\,\textsuperscript{1,2}\quad %
\href{https://orcid.org/0000-0001-6075-8081}{Andrew\,Gonzalez}%
%
\,\textsuperscript{1,2}

\textsuperscript{1}\,McGill University\quad \textsuperscript{2}\,Québec
Centre for Biodiversity Sciences\quad \textsuperscript{3}\,Université de
Montréal

\textbf{Correspondance to:}\\
Michael D. Catchen --- \texttt{michael.catchen@mcgill.ca}\\

\vfill
This work is released by its authors under a CC-BY 4.0 license\hfill\ccby\\
Last revision: \emph{\today}

\clearpage
\thispagestyle{empty}

\vfill
Ecosystems are composed of networks of interacting species. These
interactions allow communities of species to persist through time
through both neutral and adaptive processes. Still a robust
understanding of, and ability to predict and forecast interactions among
species and their consequences for network structure remains elusive.
This knowledge-gap is largely driven by a shortfall of data---although
species occurrence data has rapidly increased in the last decade,
species interaction data has not kept pace, largely due to the intrinsic
difficulty and effort required to sample interactions. These sampling
challenges bias data and hinder inferences about the structure and
dynamics of interactions networks. Here, we demonstrate the realized
false-negative rate (the percentage of species that actually interact
but for which we do not yet have a record) can be quite high, even in
thoroughly sampled systems, due to the intrinsic variation in abundances
across species in a community. We demonstrate how a null model of
occurrence detection can be used to estimate the false-negative rate in
a given dataset. One hypothesis is that interactions between ``rare''
species are themselves rare because these species are less likely to
encounter one-another than species of higher relative abundance.
However, we demonstrate that across several datasets of spatial or
temporally replicated networks, there are positive associations that
suggest these interactions actually exist but just are not observed.
Finally, we assess how false negatives influence various models of
network prediction, and recommend directly accounting for observation
error in predictive models. We conclude by discussing how the
understanding of false-negatives can inform how we design monitoring
schemes for species interactions.

\vfill

\clearpage
\linenumbers
\pagestyle{normal}

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

Species interactions drive many processes in evolutionary biology and
community ecology. A better understanding of interactions among species
is an imperative to both mitigate the potentially harmful impacts of
anthropogenic change on Earth's biodiversity (Makiola \emph{et al.}
2020) and to predict zoonotic spillover of disease to prevent future
pandemics (Becker \emph{et al.} 2021). However, meeting these challenges
is difficult because interactions are intrinsically hard to sample
(Jordano 2016). Over the past few decades biodiversity data has become
increasingly available---for example, remote-sensing has enabled
collection of data on spatial scales and resolutions previously
unimaginable (Stephenson 2020), while the adoption of open data
practices (Kenall \emph{et al.} 2014) have substantially increased the
amount of data available to ecologists. Still, widespread data about
species interactions remains elusive (Poisot \emph{et al.} 2021).
Observing an interaction between two species often requires human
observation because remote sampling methods can primarily detect
co-occurrence (but see Niedballa \emph{et al.} (2019)), and this itself
is not necessarily indicative of an interaction (Blanchet \emph{et al.}
2020). This constraint induces biases on species interaction data
subject to the spatial and temporal scales that current observation
methods can feasibly sample. This is further compounded by semantic
confusion around the word ``interaction''---for example one might
consider competition a type of species interaction, even though it is
marked by a lack of co-occurrence between species, unlike other types of
interactions, like trophism or pollination, which require both species
to be together at the same place and time. We define interaction in the
latter sense, where two species have fitness consequences on one-another
if they are in the sample place at the same time. In addition, here we
only consider direct (not higher-order) interactions.

The importance of \emph{sampling effort} and its impact on resulting
ecological data has produced a rich body of literature. The recorded
number of species in a dataset or sample depends on the total number of
observations (Walther \emph{et al.} 1995; Willott 2001)---as do
estimates of population abundance (Griffiths 1998)---in addition to
spatial coverage and species detectability. This has motivated more
quantitatively robust approaches to account for error in sampling data
in many contexts: to determine if a given species is extinct (Boakes
\emph{et al.} 2015), to determine sampling design (Moore \& McCarthy
2016),and to measure species richness across large scales (Carlson
\emph{et al.} 2020). In the context of interactions, an initial concern
was the compounding effects of limited sampling effort combined with the
amalgamation of data (across both study sites, time of year, and
taxonomic scales) could lead any empirical set of observations to
inadequately reflect the reality of how species interact (Paine 1988) or
the structure of the network as a whole (McLeod \emph{et al.} 2021).
Martinez \emph{et al.} (1999) showed that in a plant-endophyte trophic
network, network connectance is robust to sampling effort, but this was
done in the context of a system for which observation of 62,000 total
interactions derived from 164,000 plant-stems was feasible. In some
systems (e.g.~megafauna food-webs) this many observations is either
impractical or infeasible due to the absolute abundance of the species
in question.

We cannot feasibly observe all (or even most) of the direct interactions
that occur in an ecosystem. This means we can be confident two species
actually interact if we have a record of it (given an estimate of
species misidentification probability), but not at all confident that a
pair of species \emph{do not} interact if we have no record of those
species observed together. In other words, it is difficult to
distinguish \emph{true-negatives} (two species never interact) from
\emph{false-negatives} (two species interact sometimes, but we do not
have a record of it). For a concrete example of a false-negative in a
food web, see fig.~\ref{fig:concept}. Because even the most highly
sampled systems will still contain missing interactions, there is
increasing interest in combining species-level data (e.g.~traits,
abundance, range, phylogenetic relatedness, etc.) to build models to
predict interactions between species we haven't observed together before
(Strydom \emph{et al.} 2021). However, the noise of false-negatives
could impact the efficacy of our predictive models and have practical
consequences for answering questions about interactions (de Aguiar
\emph{et al.} 2019). This data constraint is amplified as the
interaction data we have is geographically biased toward the usual
suspects (Poisot \emph{et al.} 2021). We therefore need a systematic
approach to assessing these biases in the observation process and the
consequences this has for our understanding of interaction networks.

The intrinsic properties of ecological communities create several
challenges for sampling: first, species are not observed with equal
probability---we are much more likely to observe a species of high
abundance than one of very low abundance (Poisot \emph{et al.} 2015).
Canard \emph{et al.} (2012) presents a null model of food-web structure
where species encounter one-another directly in proportion to each
species' relative-abundance. This assumes that there are no associations
in species co-occurrence due to an interaction (perhaps because this
interaction is ``important'' for both species; Cazelles \emph{et al.}
(2016)), but in this paper we later show increasing strength of
associations leads to increasing probability of false-negatives in
interaction data, and that these positive associations are rampant in
existing network data. Second, observed co-occurrence is often equated
with meaningful interaction strength, but this is not necessarily the
case (Blanchet \emph{et al.} 2020)---a true ``non-interaction'' would
require that neither of two species, regardless of whether they
co-occur, exhibit any meaningful effect on the fitness of the other. So,
although co-occurrence is not directly indicative of an interaction, it
is a precondition for an interaction. Therefore observations of
``non-interactions'' between pairs of species that are outside of the
union of both species ranges do not provide any information about that
interaction, i.e.~they should be excluded from consideration.

Here, we illustrate how our confidence that a pair of species we believe
to not interact highly depends on sampling effort, and suggest that
surveys of species interactions can benefit from simulation modeling of
detection probability. We demonstrate that the realized false-negative
rate of interactions is directly related to the relative abundance of
the species involved, and demonstrate how simulation can be used to
produce a null estimate of the false-negative rate as a function of
total sampling effort (the total count of all individuals of all species
seen). We show that positive associations in co-occurrence data can
increase realized probability of false-negatives, and demonstrate these
positive associations are ubiquitous in network datasets. We conclude by
recommending that the simulation of sampling effort and species
occurrence can and should be used to help design surveys of species
interaction diversity (Moore \& McCarthy 2016), and by advocating use of
null models like those presented here as a tool for both guiding design
of surveys of species interactions and for modeling detection error in
predictive models.

\begin{figure}
\hypertarget{fig:concept}{%
\centering
\includegraphics{./figures/concept.png}
\caption{This conceptual example considers a sample of the trophic
community of bears, wolves, salmon (pink fish), pike (yellow fish),
berry trees, and aspen trees. The true metaweb (all realized
interactions across the entire spatial extent) is shown on the left. In
the center is what a hypothetical ecologist samples at each site. Notice
that although bears are observed co-occurring with both salmon and pike,
there was never a direct observation of bears eating pike, even though
they actually do. Therefore, this interaction between bears and pike is
a false negative.}\label{fig:concept}
}
\end{figure}

\hypertarget{how-many-observations-of-a-non-interaction-do-we-need-to-be-confident-its-a-true-negative}{%
\section{How many observations of a non-interaction do we need to be
confident it's a true
negative?}\label{how-many-observations-of-a-non-interaction-do-we-need-to-be-confident-its-a-true-negative}}

We start with a naive model of interaction detection: we assume that
every interacting pair of species is incorrectly observed as
not-interacting with an independent and fixed probability, which we
denote pfn and subsequently refer to as the False-Negative Rate (FNR).
If we observe the same species not-interacting \(N\) times, then the
probability of a true-negative (denoted \(p_{tn}\)) is given by
\(p_{tn}=1-(p_{fn})^N\). This relation (the probability-mass-function of
geometric distribution, a special case of the negative-binomial
distribution) is shown in Figure 1(A) for varying values of \(p_{fn}\)
and illustrates a fundamental link between our ability to reliably say
an interaction doesn't exist---\(p_{tn}\)---and the number of times we
have observed a given species. In addition, note that also there is no
non-zero \(p_{fn}\) for which we can ever prove that an interaction does
not exist---no matter how many observations of non-interactions \(N\) we
have, \(p_{tn}<1\).

From fig.~\ref{fig:geometric}(A) it is clear that the more often we see
two species co-occurring, but not interacting, the more likely the
interaction is a true negative. This has several practical consequences:
first it means negatives taken outside the overlap of the range of each
species aren't informative. Second, we can use this relation to compute
the expected number of total observations needed to obtain a ``goal''
number of observations of a particular pair of species
(fig.~\ref{fig:geometric}(B)). As an example, if we hypothesize that
\(A\) and \(B\) do not interact, and we want to see species \(A\) and
\(B\) both co-occurring and not interacting 10 times to be confident
this is a true negative, then we need an expected 1000 observations of
all species if the relative abundances of \(A\) and \(B\) are both
\(0.1\).

Because the true FNR is latent, we can never actually be sure what the
actual number of false negatives in our data---however, we can use
simulation to estimate it for datasets of a given size using neutral
models of observation. If some of the ``worst-case'' FNRs presented in
fig.~\ref{fig:geometric}(A) seem unrealistically high, consider that
species are observed in proportion to their relative abundance. In the
next section we demonstrate that the distribution of abundance in
ecosystems can lead to very high realized values of FNR (\(p_{fn}\))
simply as an artifact of sampling effort.

\begin{figure}
\hypertarget{fig:geometric}{%
\centering
\includegraphics{./figures/fig1.png}
\caption{(A) The probability that an observed interaction is a true
negative (y-axis) given how many times it has been sampled as a
non-interaction (x-axis). Each color reflects a different value of
\(p_{fn}\), the false-negative rate (FNR)---this is effectively the cdf
of the geometric distribution. (B): The expected needed observations of
all individuals of all species (y-axis) required to obtain a goal number
of observations (colors) of a particular species, and a function of the
relative abundance of that focal species (x-axis). (C): False negative
rate (y-axis) as a function of total sampling effort (x-axis) and
network size, computed using the method described above. For 500
independent draws from the niche model (Williams \& Martinez (2000)) at
varying levels of species richness (colors) with connectance drawn
according to the flexible-links model (MacDonald \emph{et al.} (2020))
as described in the main text. For each draw from the niche model, 200
sets of 1500 observations are simulated, for which each the mean false
negative rate at each observation-step is computed. Means denoted with
points, with 1 in the first shade and 2 in the second. (D): Same as (C),
except using empirical food webs from Mangal database, where richness.
The outlier on (D) is a 714 species food-web.}\label{fig:geometric}
}
\end{figure}

\hypertarget{false-negatives-as-a-product-of-relative-abundance}{%
\section{False-negatives as a product of relative
abundance}\label{false-negatives-as-a-product-of-relative-abundance}}

We now show that the realized FNR changes drastically with sampling
effort due to the intrinsic variation of the abundance of individuals of
each species within a community. We do this by simulating the process of
observation of species interactions, applied both to 243 empirical food
webs from the Mangal database (Banville \emph{et al.} 2021) and random
food-webs generated using the niche model (Williams \& Martinez 2000), a
simple generative model of food-web structure. Our neutral model of
observation assumes each observed species is drawn from the distribution
in proportion to each species' abundance at that place and time. The
abundance distribution of a community can be reasonably-well described
by a log-normal distribution (Volkov \emph{et al.} 2003). In addition to
the log-normal distribution, we also tested the case where the abundance
distribution is derived from power-law scaling \(Z^{(log(T_i)-1)}\)
where \(T_i\) is the trophic level of species \(i\) and \(Z\) is a
scaling coefficient (Savage \emph{et al.} 2004), which yields the same
qualitative behavior. The practical consequence of abundance
distributions spanning many orders of magnitude of abundance is that
observing two ``rare'' species interacting requires two low probability
events: observing two rare species at the same time.

To simulate the process of observation, for an ecological network \(M\)
with \(S\) species, we sample abundances for each species from a
standard-log-normal distribution. For each true interaction in the
adjacency matrix \(M\) (i.e. \(M_{ij}=1\)) we estimate the probability
of observing both species \(i\) and \(j\) at a given place and time by
simulating \(n\) observations of all individuals of any species, where
the species of the individual observed at the \(\{1,2,\dots,n\}\)-th
observation is drawn from the generated log-normal distribution of
abundances. For each pair of species \((i,j)\), if both \(i\) and \(j\)
are observed within the n-observations, the interaction is tallied as a
true positive if \(M_{ij}=1\). If only one of \(i\) or \(j\) are
observed---but not both---in these \(n\) observations, but \(M_{ij}=1\),
this is counted as a false-negative, and a true-negative otherwise.

In fig.~\ref{fig:geometric}(C) we see this model of observation applied
to niche model networks across varying levels of species richness, and
in fig.~\ref{fig:geometric}(D) the observation model applied to Mangal
food webs. For all niche model simulations in this manuscript, for a
given number of species \(S\) the number of interactions is drawn from
the flexible-links model fit to Mangal data (MacDonald \emph{et al.}
2020), effectively drawing the number of interactions \(L\) for a random
niche model food-web as

\[L \sim  \text{BetaBinomial}(S^2-S+1,\mu\phi, 1-\mu\phi)\]

where the MAP estimate of \((\mu, \phi)\) applied to Mangal data from
(MacDonald \emph{et al.} 2020) is \((\mu=0.086, \phi=24.3)\). All
simulations were done with 500 independent replicates per unique number
of observations \(n\). All analyses presented here are done in Julia
v1.8 (Bezanson \emph{et al.} 2015) using both EcologicalNetworks.jl v0.5
and Mangal.jl v0.4 (Banville \emph{et al.} 2021) and are hosted at
(GITHUB\_LINK\_TODO). Note that the empirical data, for the reasons
described above, very likely already contains many false negatives,
we'll revisit this issue in the final section.

From fig.~\ref{fig:geometric}(C) it is evident that the number of
species considered in a study is inseparable from the false-negative
rate in that study, and this effect should be taken into account when
designing samples of ecological networks in the future. We see a similar
qualitative pattern in fig.~\ref{fig:geometric}(D) where the FNR drops
off quickly as a function of observation effort, mediated by total
richness. The practical consequence of the bottom row of Figure 1 is
whether the total number of observations of all species (the x-axis) for
the range of possible FNR we deem acceptable (the y-axis) is feasible.
This raises two points: first, empirical data on interactions are
subject to the practical limitations of funding and human-work hours,
and therefore existing data tend to fall on the order of hundreds or
thousands observations of individuals per site. Clear aggregation of
data on sampling effort has proven difficult to find and a meta-analysis
of network data and sampling effort seems both pertinent and necessary,
in addition to the effects of aggregation of interactions across
taxonomic scales (Gauzens \emph{et al.} 2013; Giacomuzzo \& Jordán
2021). This inherent limitation on in-situ sampling means we should
optimize where we sample across space so that for a given number of
samples, we obtain the maximum information possible. Second, what is
meant by ``acceptable'' FNR? This raises the question: does a shifting
FNR lead to rapid transitions in our ability inference and predictions
about the structure and dynamics of networks, or does it produce a
roughly linear decay in model efficacy? We explore this in the next
section.

We conclude this section by advocating for the use of neutral models
similar to above to generate expectations about the number of
false-negatives in a data set of a given size. This could prove fruitful
both for designing surveys of interactions but also because we may want
to incorporate models of imperfect detection error into predictive
interactions models, as Joseph (2020) does for species occurrence
modeling. Additionally, we emphasize that one must consider the context
for sampling---is the goal to detect a particular species (as in
fig.~\ref{fig:geometric}(C)), or to get a representative sample of
interactions across the species pool? This argument is well-considered
when sampling individual species (Willott 2001), but has not yet been
adopted for designing samples of communities.

\hypertarget{positive-associations-increase-the-false-negative-rate}{%
\section{Positive associations increase the false-negative
rate}\label{positive-associations-increase-the-false-negative-rate}}

This model above doesn't consider the possibility that there are
positive or negative associations which shift the probability of
observing two species together due to their interaction (Cazelles
\emph{et al.} 2016). However, here we demonstrate that the probability
of observing a false negative can be higher if there is some positive
association in the occurrence of species \(A\) and \(B\). If we denote
the probability that we observe the co-occurrence of two species \(A\)
and \(B\) that we know interact as \(P(AB)\) and if there is no
association between the marginal probabilities of observing \(A\) and
observing \(B\), denoted \(P(A)\) and \(P(B)\) respectively, then the
probability of observing their co-occurrence \(P(AB) = P(A)P(B)\). In
the other case where there is some positive strength of association
between observing both \(A\) and \(B\) because this interaction is
``important'' for each species, then the probability of observation both
\(A\) and \(B\), \(P(AB)\), is greater than \(P(A)P(B)\) as \(P(A)\) and
\(P(B)\) are not independent and instead are positively correlated,
i.e.~\(P(AB)> P(A)P(B)\). In this case, the probability of observing a
false negative in our naive model from fig.~\ref{fig:geometric}(A) is
\(p_{fn}= 1-P(AB)\), which due to the above inequality implies
\(p_{fn} >1-P(A)P(B)\). This indicates an increasingly greater
probability of a false negative as \(P(AB) \to P(AB) \gg P(A)P(B)\).
However, this still does not consider variation in species abundance in
space and time (Poisot \emph{et al.} 2015). If positive or negative
associations between species structure variation in the distribution of
\(P(AB)\) across space/time, then the spatial/temporal biases induced by
data collection would further impact the realized false negative rate,
as the probability of false negative would not be constant for each pair
of species across sites.

To test for this association in data we scoured Mangal for datasets with
many spatial or temporal replicates of the same system. For each
dataset, we compute the marginal probability \(P(A)\) of occurrence of
each species \(A\) across all networks in the dataset. For each pair of
interacting species \(A\) and \(B\), we then compute and compare the
probability of co-occurrence if each species occurs independently,
\(P(A)P(B)\), to the empirical joint probability of co-occurrence,
\(P(AB)\). Following our analysis above, if \(P(AB)\) is greater than
\(P(A)P(B)\), then we expect our neutral estimates of the FNR above to
underestimate the realized FNR. In fig.~\ref{fig:mangal}, we see the
difference between \(P(AB)\) and \(P(A)P(B)\) for the seven suitable
datasets with enough spatio-temporal replicates and a shared taxonomic
backbone (meaning all individual networks use common species
identifiers) found on Mangal to perform this analysis. Further details
about each dataset are reported in tbl.~\ref{tbl:id}.

\begin{figure}
\hypertarget{fig:mangal}{%
\centering
\includegraphics{./figures/fig2.png}
\caption{The difference between joint-probability of co-occurrence
(\(P(AB)\)) and expected probability of co-occurrence under independence
(\(P(A)P(B)\)) for interacting species for each dataset. The red-dashed
line indicates 0. Each histogram represents a density, meaning the area
of the entire curve sums to 1. The continuous density estimate (computed
using local smoothing) is shown in grey. The p-value on each plot is the
result of a one-sided t-test comparing the mean of each distribution to
0.}\label{fig:mangal}
}
\end{figure}

\hypertarget{tbl:id}{}
\begin{longtable}[]{@{}ccccccccc@{}}
\caption{\label{tbl:id}This table describes the datasets used in the
above analysis (Fig 2). The table reports the type of each dataset, the
total number of networks in each dataset \((N)\), the total species
richness in each dataset \((S)\), the connectance of each metaweb (all
interactions across the entire spatial-temporal extent) \((C)\), the
mean species richness across each local network \(S\), the mean
connectance of each local network \(C\), the mean \(\beta\)-diversity
among overlapping species across all pairs of network species
(\(\bar{\beta}_{OS}\)), and the mean \(\beta\)-diversity among all
species in the metaweb (\(\bar{\beta}_{WN}\)). Both metrics are computed
using KGL \(\beta\)-diversity (Koleff \emph{et al.}
2003)}\tabularnewline
\toprule
Network & Type & \(N\) & \(S\) & \(C\) & \(\bar{S}\) & \(\bar{C}\) &
\(\bar{\beta}_{OS}\) & \(\bar{\beta}_{WN}\)\tabularnewline
\midrule
\endfirsthead
\toprule
Network & Type & \(N\) & \(S\) & \(C\) & \(\bar{S}\) & \(\bar{C}\) &
\(\bar{\beta}_{OS}\) & \(\bar{\beta}_{WN}\)\tabularnewline
\midrule
\endhead
Kopelke \emph{et al.} (2017) & Food Web & 100 & 98 & 0.037 & 7.87 &
0.142 & 1.383 & 1.972\tabularnewline
Thompson \& Townsend (2000) & Food Web & 18 & 566 & 0.014 & 80.67 &
0.049 & 1.617 & 1.594\tabularnewline
Havens (1992) & Food Web & 50 & 188 & 0.065 & 33.58 & 0.099 & 1.468 &
1.881\tabularnewline
Ponisio \emph{et al.} (2017) & Pollinator & 100 & 226 & 0.079 & 23.0 &
0.056 & 1.436 & 1.870\tabularnewline
Hadfield \emph{et al.} (2014) & Host-Parasite & 51 & 327 & 0.085 & 32.71
& 0.337 & 1.477 & 1.952\tabularnewline
Closs \& Lake (1994) & Food Web & 12 & 61 & 0.14 & 29.09 & 0.080 & 1.736
& 1.864\tabularnewline
CaraDonna \emph{et al.} (2017) & Pollinator & 86 & 122 & 0.18 & 21.42 &
0.312 & 1.527 & 1.907\tabularnewline
\bottomrule
\end{longtable}

In each of these datasets, the joint probably of co-occurrence \(P(AB)\)
is decisively greater than our expectation if species co-occur in
proportion to their relative abundance \(P(A)P(B)\). This suggests that
there may not be as many ``neutrally forbidden links'' (Canard \emph{et
al.} 2012) as we might think, and that the reason we do not have records
of interactions between rare species is probably due to observation
error. This has serious ramifications for the widely observed property
of nestedness seen in bipartite networks (Bascompte \& Jordano
2007)---perhaps the reason we have lots of observations between
generalists is because they are more abundant, and this is particularly
relevant as we have strong evidence that generalism drives abundance
(Song \emph{et al.} 2022), not vice-versa.

\hypertarget{the-impact-of-false-negatives-on-network-properties-and-prediction}{%
\section{The impact of false-negatives on network properties and
prediction}\label{the-impact-of-false-negatives-on-network-properties-and-prediction}}

Here, we assess the effect of false negatives on our ability to make
predictions about interactions, as well as their effect on network
structure. The prevalence of false-negatives in our data is the catalyst
for interaction prediction in the first place, and as a result methods
have been proposed to counteract this bias Poisot \emph{et al.} (2022).
However, it is feasible that the FNR is so high that it could induce too
much noise for an interaction prediction model to detect the signal of
possible interaction between species.

To test this we use the dataset from Hadfield \emph{et al.} (2014) that
describes host-parasite interaction networks sampled across 51 sites,
and the same method as Strydom \emph{et al.} (2021) to extract latent
features for each species in this dataset based on co-occurrence. We
then predict a metaweb (equivalent to predicting true or false for an
interaction for each species pair, effectively a binary classification
problem) from these species-level features using four candidate models
for binary classification---three often used machine-learning (ML)
methods (Boosted Regression Tree (BRT), Random Forest (RF), Decision
Tree (DT)), and one naive model from classic statistics (Logistic
Regression (LR)). Each of the ML models are bootstrap aggregated (or
bagged) with 100 replicates each. We partition the data into 80-20
training-test split, and then seed the training data with false
negatives at varying rates, but crucially do nothing to the test data.
We fit all of these models using MLJ.jl, a high-level Julia framework
for a wide-variety of ML models (Blaom \emph{et al.} 2020). We evaluate
the efficacy of these models using two common measures of binary
classifier performance: the area under the receiver-operator curve
(ROC-AUC) and the area under the precision-recall curve (PR-AUC), for
more details see (Poisot 2022) Here, PR-AUC is slightly more relevant as
it is a better indicator of prediction of false-negatives. The results
of these simulations are shown in fig.~\ref{fig:addedfnr}(A) and (B).

\begin{figure}
\hypertarget{fig:addedfnr}{%
\centering
\includegraphics{./figures/fig3.png}
\caption{(A) The area-under the receiver-operator curve (ROC-AUC) and
(B) The area-under the precision-recall curve (PR-AUC; right) for each
different predictive model (colors/shapes) across a spectrum of the
proportion of added false negatives (x-axis). (C) The mean trophic-level
of all species in a network generated with the niche model across
different species richnesses (colors). For each value of the FNR, the
mean trophic level was computed across 50 replicates. The shaded region
for each line is one standard-deviation across those
replicates.}\label{fig:addedfnr}
}
\end{figure}

One interesting result seen in fig.~\ref{fig:addedfnr}(A) and (B) is
that the ROC-AUC value does not approach random in the same way the
PR-AUC curve does as we increase the added FNR. The reason for this is
that ROC-AUC is fundamentally not as useful a metric in assessing
predictive capacity as PR-AUC. As we keep adding more false-negatives,
the network eventually becomes a zeros matrix, and these models can
still learn to predict ``no-interaction'' for all possible species
pairs, which does far better than random guessing (ROC-AUC
\textgreater{} 0.5) in terms of the true and false positive rates (the
components of ROC-AUC). This highlights a more broad issue of label
class imbalance, meaning there are far more non-interactions than
interactions in data. A full treatment of the importance of
class-balance is outside the scope of this paper, but is explored
in-depth in (Poisot 2022).

Although these ML models are surprisingly performant at link prediction
given their simplicity, there have been several major developments in
the field applying deep-learning methods to many tasks in network
inference and prediction---namely graph-representation learning (GRL,
Khoshraftar \& An (2022)) and graph convolutional networks (Zhang
\emph{et al.} 2019). At this time, these advances can not yet be applied
to ecological networks because they require far more data than we
currently have. We already have lots of features that could be used as
inputs into these models (i.e.~species level data about occurrence,
genomes, abundance, etc.), but our network datasets barely get into the
hundreds of local networks sampled across space and time
tbl.~\ref{tbl:id}. Once we start to get into the thousands, these models
will become more useful, but this can only be done with systematic
monitoring of interactions. This again highlights the need to optimize
our sampling effort to maximize the amount of information contained in
our data given the expensive nature of sampling interactions.

We also consider how the FNR affects network properties. In
fig.~\ref{fig:addedfnr}(C) we see the mean trophic level across networks
simulated using the niche model (as above), across a spectrum of FNR
values. In addition to the clear dependence on richness, we see that
mean trophic level, despite varying widely between niche model
simulations, tends to be relatively robust to false negatives and does
not deviate widely from the true value until very large FNRs,
i.e.~\(p_{fn} > 0.7\).

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

Species interactions enable the persistence and functioning of
ecosystems, but our understanding of interactions is limited due to the
intrinsic difficulty of sampling. Here we have provided a null model for
the expected number of false-negatives in an interaction dataset. We
demonstrated that we expect many false-negatives in species interaction
datasets purely due to the intrinsic variation of abundances within a
community. We also, for the first time to our knowledge, measured the
strength of association between co-occurrence and interactions (Cazelles
\emph{et al.} 2016) across many empirical systems for the first time,
and found that these positive associations are both very common, and
showed algebraically that they increase the realized FNR. We have also
shown that false-negatives could further impact our ability to both
predict interactions and infer properties of the networks, which
highlights the need for further research into methods for correcting
this bias in existing data.

A better understanding of how false-negatives impact our inference of
network structure and dynamics is prediction of ecological networks by
both using information about individual species interactions and the
structure of metawebs on large scales is a practical necessity.
False-negatives could pose a problem for many forms of inference in
network ecology. For example, inferring the structural or dynamic
stability of a network could be prone to error if the observed network
is not sampled ``enough.'' What exactly ``enough'' means is then
specific to the application, and should be assessed via methods like
those here when designing samples. Further, predictions about network
rewiring (Thompson \& Gonzalez 2017) due to range shifts in response to
climate change could be error-prone without accounting for interactions
that have not been observed but that still may become climatically
infeasible. As is evident from Figure 1(A), we can never guarantee there
are no false-negatives in data. In recent years, there has been interest
toward explicitly accounting for false-negatives in models (Stock
\emph{et al.} 2017; Young \emph{et al.} 2021), and a predictive approach
to networks---rather than expecting our samples to fully capture all
interactions (Strydom \emph{et al.} 2021). As a result, better models
for predicting interactions are needed for interaction networks. This
includes explicitly accounting for observation error (Johnson \&
Larremore 2021)---certain classes of models have been used to reflect
hidden states which account for detection error in occupancy modeling
(Joseph 2020), and could be integrated in the predictive models of
interactions in the future.

A brief caveat here is that we do not consider the rate of
false-positives---in large part false-positives can be explained by
misidentification of species, although this could be a relevant
consideration in some cases. The same logic that we apply to
false-negatives could easily be applied to false-positives, e.g.~that we
can be much more confident that an interaction is a true positive if we
have observed it 50 times rather than only once, and we could similarly
model this using the geometric distribution as in
fig.~\ref{fig:geometric}(A). However, because ecological networks are so
sparse, there are far more negatives than positives in the dataset, and
therefore likely to be far more false-negatives than false-positives in
absolute terms.

This work has several practical consequences for the design of
interaction samples. Simulating the process of observation could be a
powerful tool for estimating the sampling effort required by a study
that takes relative abundance into account, and provides a null baseline
for expected FNR. It is necessary to take the size of the species pool
into account when deciding how many total samples is sufficient for an
``acceptable'' FNR (fig.~\ref{fig:geometric}(C \& D)). Further the
spatial and temporal turnover of interactions means any approach to
sampling prioritization must be spatiotemporal. We demonstrated earlier
that observed negatives outside of the range of both species aren't
informative, and therefore using species distribution models could aid
in this spatial prioritization of sampling sites.

Our work highlights the need for a quantitatively robust approach to
sampling design, both for interactions (Jordano 2016) and all other
aspects of biodiversity (Carlson \emph{et al.} 2020). As anthropogenic
forces create rapid shifts in our planet's climate and biosphere, this
is an imperative to maximize the amount of ecological information we get
in our finite samples, and make our inferences and decisions based on
this data as robust as possible. Where we choose to sample, and how
often we choose to sample there, has strong impacts on the inferences we
make from data. Incorporating a better understanding of sampling effort
and bias to the design of biodiversity monitoring systems, and the
inference and predictive models we apply to this data, is imperative in
understanding how biodiversity is changing, and making actionable
forecasts about the future of ecological interactions on our planet.

\hypertarget{references}{%
\section*{References}\label{references}}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\hypertarget{ref-Banville2021ManJl}{}%
Banville, F., Vissault, S. \& Poisot, T. (2021). Mangal.jl and
EcologicalNetworks.jl: Two complementary packages for analyzing
ecological networks in Julia. \emph{Journal of Open Source Software}, 6,
2721.

\leavevmode\hypertarget{ref-Bascompte2007PlaMut}{}%
Bascompte, J. \& Jordano, P. (2007). Plant-Animal Mutualistic Networks:
The Architecture of Biodiversity. \emph{Annual Review of Ecology,
Evolution, and Systematics}, 38, 567--593.

\leavevmode\hypertarget{ref-Becker2021OptPre}{}%
Becker, D.J., Albery, G.F., Sjodin, A.R., Poisot, T., Bergner, L.M.,
Dallas, T.A., \emph{et al.} (2021). Optimizing predictive models to
prioritize viral discovery in zoonotic reservoirs.

\leavevmode\hypertarget{ref-Bezanson2015JulFre}{}%
Bezanson, J., Edelman, A., Karpinski, S. \& Shah, V.B. (2015). Julia: A
Fresh Approach to Numerical Computing.

\leavevmode\hypertarget{ref-Blanchet2020CooNot}{}%
Blanchet, F.G., Cazelles, K. \& Gravel, D. (2020). Co-occurrence is not
evidence of ecological interactions. \emph{Ecology Letters}, 23,
1050--1063.

\leavevmode\hypertarget{ref-Blaom2020MljJul}{}%
Blaom, A.D., Kiraly, F., Lienart, T., Simillides, Y., Arenas, D. \&
Vollmer, S.J. (2020). MLJ: A Julia package for composable machine
learning. \emph{Journal of Open Source Software}, 5, 2704.

\leavevmode\hypertarget{ref-Boakes2015InfSpe}{}%
Boakes, E.H., Rout, T.M. \& Collen, B. (2015). Inferring species
extinction: The use of sighting records. \emph{Methods in Ecology and
Evolution}, 6, 678--687.

\leavevmode\hypertarget{ref-Canard2012EmeStr}{}%
Canard, E., Mouquet, N., Marescot, L., Gaston, K.J., Gravel, D. \&
Mouillot, D. (2012). Emergence of Structural Patterns in Neutral Trophic
Networks. \emph{PLOS ONE}, 7, e38295.

\leavevmode\hypertarget{ref-CaraDonna2017IntRew}{}%
CaraDonna, P.J., Petry, W.K., Brennan, R.M., Cunningham, J.L.,
Bronstein, J.L., Waser, N.M., \emph{et al.} (2017). Interaction rewiring
and the rapid turnover of plantpollinator networks. \emph{Ecology
Letters}, 20, 385--394.

\leavevmode\hypertarget{ref-Carlson2020WhaWou}{}%
Carlson, C.J., Dallas, T.A., Alexander, L.W., Phelan, A.L. \& Phillips,
A.J. (2020). What would it take to describe the global diversity of
parasites? \emph{Proceedings of the Royal Society B: Biological
Sciences}, 287, 20201841.

\leavevmode\hypertarget{ref-Cazelles2016TheSpe}{}%
Cazelles, K., Araújo, M.B., Mouquet, N. \& Gravel, D. (2016). A theory
for species co-occurrence in interaction networks. \emph{Theoretical
Ecology}, 9, 39--48.

\leavevmode\hypertarget{ref-Closs1994SpaTem}{}%
Closs, G.P. \& Lake, P.S. (1994). Spatial and Temporal Variation in the
Structure of an Intermittent-Stream Food Web. \emph{Ecological
Monographs}, 64, 1--21.

\leavevmode\hypertarget{ref-deAguiar2019RevBia}{}%
de Aguiar, M.A.M., Newman, E.A., Pires, M.M., Yeakel, J.D., Boettiger,
C., Burkle, L.A., \emph{et al.} (2019). Revealing biases in the sampling
of ecological interaction networks. \emph{PeerJ}, 7, e7566.

\leavevmode\hypertarget{ref-Gauzens2013FooAgg}{}%
Gauzens, B., Legendre, S., Lazzaro, X. \& Lacroix, G. (2013). Food-web
aggregation, methodological and functional issues. \emph{Oikos}, 122,
1606--1615.

\leavevmode\hypertarget{ref-Giacomuzzo2021FooWeb}{}%
Giacomuzzo, E. \& Jordán, F. (2021). Food web aggregation: Effects on
key positions. \emph{Oikos}, 130, 2170--2181.

\leavevmode\hypertarget{ref-Griffiths1998SamEff}{}%
Griffiths, D. (1998). Sampling effort, regression method, and the shape
and slope of sizeabundance relations. \emph{Journal of Animal Ecology},
67, 795--804.

\leavevmode\hypertarget{ref-Hadfield2014TalTwo}{}%
Hadfield, J.D., Krasnov, B.R., Poulin, R. \& Nakagawa, S. (2014). A Tale
of Two Phylogenies: Comparative Analyses of Ecological Interactions.
\emph{The American Naturalist}, 183, 174--187.

\leavevmode\hypertarget{ref-Havens1992ScaStr}{}%
Havens, K. (1992). Scale and Structure in Natural Food Webs.
\emph{Science}, 257, 1107--1109.

\leavevmode\hypertarget{ref-Johnson2021BayEst}{}%
Johnson, E.K. \& Larremore, D.B. (2021). Bayesian estimation of
population size and overlap from random subsamples.

\leavevmode\hypertarget{ref-Jordano2016SamNet}{}%
Jordano, P. (2016). Sampling networks of ecological interactions.
\emph{Functional Ecology}, 30, 1883--1893.

\leavevmode\hypertarget{ref-Joseph2020NeuHie}{}%
Joseph, M.B. (2020). Neural hierarchical models of ecological
populations. \emph{Ecology Letters}, 23, 734--747.

\leavevmode\hypertarget{ref-Kenall2014OpeFut}{}%
Kenall, A., Harold, S. \& Foote, C. (2014). An open future for
ecological and evolutionary data? \emph{BMC Evolutionary Biology}, 14,
66.

\leavevmode\hypertarget{ref-Khoshraftar2022SurGra}{}%
Khoshraftar, S. \& An, A. (2022). A Survey on Graph Representation
Learning Methods.

\leavevmode\hypertarget{ref-Koleff2003MeaBet}{}%
Koleff, P., Gaston, K.J. \& Lennon, J.J. (2003). Measuring beta
diversity for presenceabsence data. \emph{Journal of Animal Ecology},
72, 367--382.

\leavevmode\hypertarget{ref-Kopelke2017FooStr}{}%
Kopelke, J.-P., Nyman, T., Cazelles, K., Gravel, D., Vissault, S. \&
Roslin, T. (2017). Food-web structure of willow-galling sawflies and
their natural enemies across Europe. \emph{Ecology}, 98, 1730--1730.

\leavevmode\hypertarget{ref-MacDonald2020RevLin}{}%
MacDonald, A.A.M., Banville, F. \& Poisot, T. (2020). Revisiting the
Links-Species Scaling Relationship in Food Webs. \emph{Patterns}, 1.

\leavevmode\hypertarget{ref-Makiola2020KeyQue}{}%
Makiola, A., Compson, Z.G., Baird, D.J., Barnes, M.A., Boerlijst, S.P.,
Bouchez, A., \emph{et al.} (2020). Key Questions for Next-Generation
Biomonitoring. \emph{Frontiers in Environmental Science}, 7.

\leavevmode\hypertarget{ref-Martinez1999EffSam}{}%
Martinez, N.D., Hawkins, B.A., Dawah, H.A. \& Feifarek, B.P. (1999).
Effects of Sampling Effort on Characterization of Food-Web Structure.
\emph{Ecology}, 80, 1044--1055.

\leavevmode\hypertarget{ref-McLeod2021SamAsy}{}%
McLeod, A., Leroux, S.J., Gravel, D., Chu, C., Cirtwill, A.R., Fortin,
M.-J., \emph{et al.} (2021). Sampling and asymptotic network properties
of spatial multi-trophic networks. \emph{Oikos}, 130, 2250--2259.

\leavevmode\hypertarget{ref-Moore2016OptEco}{}%
Moore, A.L. \& McCarthy, M.A. (2016). Optimizing ecological survey
effort over space and time. \emph{Methods in Ecology and Evolution}, 7,
891--899.

\leavevmode\hypertarget{ref-Niedballa2019AssAna}{}%
Niedballa, J., Wilting, A., Sollmann, R., Hofer, H. \& Courtiol, A.
(2019). Assessing analytical methods for detecting spatiotemporal
interactions between species from camera trapping data. \emph{Remote
Sensing in Ecology and Conservation}, 5, 272--285.

\leavevmode\hypertarget{ref-Paine1988RoaMap}{}%
Paine, R.T. (1988). Road Maps of Interactions or Grist for Theoretical
Development? \emph{Ecology}, 69, 1648--1654.

\leavevmode\hypertarget{ref-Poisot2022GuiPre}{}%
Poisot, T. (2022). Guidelines for the prediction of species interactions
through binary classification.

\leavevmode\hypertarget{ref-Poisot2021GloKno}{}%
Poisot, T., Bergeron, G., Cazelles, K., Dallas, T., Gravel, D.,
MacDonald, A., \emph{et al.} (2021). Global knowledge gaps in species
interaction networks data. \emph{Journal of Biogeography}, 48,
1552--1563.

\leavevmode\hypertarget{ref-Poisot2022NetEmb}{}%
Poisot, T., Ouellet, M.-A., Mollentze, N., Farrell, M.J., Becker, D.J.,
Brierly, L., \emph{et al.} (2022). Network embedding unveils the hidden
interactions in the mammalian virome.

\leavevmode\hypertarget{ref-Poisot2015SpeWhy}{}%
Poisot, T., Stouffer, D.B. \& Gravel, D. (2015). Beyond species: Why
ecological interaction networks vary through space and time.
\emph{Oikos}, 124, 243--251.

\leavevmode\hypertarget{ref-Ponisio2017OppAtt}{}%
Ponisio, L.C., Gaiarsa, M.P. \& Kremen, C. (2017). Opportunistic
attachment assembles plantpollinator networks. \emph{Ecology Letters},
20, 1261--1272.

\leavevmode\hypertarget{ref-Savage2004EffBod}{}%
Savage, V.M., Gillooly, J.F., Brown, J.H., West, G.B. \& Charnov, E.L.
(2004). Effects of Body Size and Temperature on Population Growth.
\emph{The American Naturalist}, 163, 429--441.

\leavevmode\hypertarget{ref-Song2022GenDri}{}%
Song, C., Simmons, B.I., Fortin, M.-J. \& Gonzalez, A. (2022).
Generalism drives abundance: A computational causal discovery approach.
\emph{PLOS Computational Biology}, 18, e1010302.

\leavevmode\hypertarget{ref-Stephenson2020TecAdv}{}%
Stephenson, P. (2020). Technological advances in biodiversity
monitoring: Applicability, opportunities and challenges. \emph{Current
Opinion in Environmental Sustainability}, Open issue 2020 part A:
Technology Innovations and Environmental Sustainability in the
Anthropocene, 45, 36--41.

\leavevmode\hypertarget{ref-Stock2017LinFil}{}%
Stock, M., Poisot, T., Waegeman, W. \& De Baets, B. (2017). Linear
filtering reveals false negatives in species interaction data.
\emph{Scientific Reports}, 7, 45908.

\leavevmode\hypertarget{ref-Strydom2021RoaPre}{}%
Strydom, T., Catchen, M.D., Banville, F., Caron, D., Dansereau, G.,
Desjardins-Proulx, P., \emph{et al.} (2021). A roadmap towards
predicting species interaction networks (across space and time).
\emph{Philosophical Transactions of the Royal Society B: Biological
Sciences}, 376, 20210063.

\leavevmode\hypertarget{ref-Thompson2017DisGov}{}%
Thompson, P.L. \& Gonzalez, A. (2017). Dispersal governs the
reorganization of ecological networks under environmental change.
\emph{Nature Ecology \& Evolution}, 1, 1--8.

\leavevmode\hypertarget{ref-Thompson2000ResSol}{}%
Thompson, R.M. \& Townsend, C.R. (2000). Is resolution the solution?:
The effect of taxonomic resolution on the calculated properties of three
stream food webs. \emph{Freshwater Biology}, 44, 413--422.

\leavevmode\hypertarget{ref-Volkov2003NeuThe}{}%
Volkov, I., Banavar, J.R., Hubbell, S.P. \& Maritan, A. (2003). Neutral
theory and relative species abundance in ecology. \emph{Nature}, 424,
1035--1037.

\leavevmode\hypertarget{ref-Walther1995SamEff}{}%
Walther, B.A., Cotgreave, P., Price, R.D., Gregory, R.D. \& Clayton,
D.H. (1995). Sampling Effort and Parasite Species Richness.
\emph{Parasitology Today}, 11, 306--310.

\leavevmode\hypertarget{ref-Williams2000SimRul}{}%
Williams, R.J. \& Martinez, N.D. (2000). Simple rules yield complex food
webs. \emph{Nature}, 404, 180--183.

\leavevmode\hypertarget{ref-Willott2001SpeAcc}{}%
Willott, S.j. (2001). Species accumulation curves and the measure of
sampling effort. \emph{Journal of Applied Ecology}, 38, 484--486.

\leavevmode\hypertarget{ref-Young2021RecPla}{}%
Young, J.-G., Valdovinos, F.S. \& Newman, M.E.J. (2021). Reconstruction
of plantpollinator networks from observational data. \emph{Nature
Communications}, 12, 3911.

\leavevmode\hypertarget{ref-Zhang2019GraCon}{}%
Zhang, S., Tong, H., Xu, J. \& Maciejewski, R. (2019). Graph
convolutional networks: A comprehensive review. \emph{Computational
Social Networks}, 6, 11.

\end{CSLReferences}

\end{document}