### 2022-10-16 draft.tex
Deploying to gh-pages from @ gottacatchenall/ms_false_negatives@e411d6fd2bc3a08964824a43ec81f379ae7c2be2 🚀

%!TEX TS-program = xelatex
\documentclass[11pt]{article}

\usepackage[english]{babel}

\usepackage{amsmath,amssymb,amsfonts}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{stix2}
\usepackage[scaled]{helvet}
\usepackage[scaled]{inconsolata}

\usepackage{lastpage}

\usepackage{setspace}

\usepackage{ccicons}

\usepackage[hang,flushmargin]{footmisc}

\usepackage{geometry}

\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}

\usepackage{fancyhdr}
\renewcommand{\headrulewidth}{0pt}\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

\makeatletter
\newcounter{tableno}
\newenvironment{tablenos:no-prefix-table-caption}{
  \caption@ifcompatibility{}{
    \let\oldthetable\thetable
    \let\oldtheHtable\theHtable
    \renewcommand{\thetable}{tableno:\thetableno}
    \renewcommand{\theHtable}{tableno:\thetableno}
    \stepcounter{tableno}
    \captionsetup{labelformat=empty}
  }
}{
  \caption@ifcompatibility{}{
    \captionsetup{labelformat=default}
    \let\thetable\oldthetable
    \let\theHtable\oldtheHtable
    \addtocounter{table}{-1}
  }
}
\makeatother

\usepackage{array}
\newcommand{\PreserveBackslash}[1]{\let\temp=\\#1\let\\=\temp}
\let\PBS=\PreserveBackslash

\usepackage[breaklinks=true]{hyperref}
\hypersetup{colorlinks,%
citecolor=blue,%
filecolor=blue,%
linkcolor=blue,%
urlcolor=blue}
\usepackage{url}

\usepackage{caption}
\setcounter{secnumdepth}{0}
\usepackage{cleveref}

\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth
\else\Gin@nat@width\fi}
\makeatother
\let\Oldincludegraphics\includegraphics
\renewcommand{\includegraphics}[1]{\Oldincludegraphics[width=\maxwidth]{#1}}

\usepackage{longtable}
\usepackage{booktabs}

\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}

\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[3] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1 \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces\fi
  % set entry spacing
  \ifnum #2 > 0
  \setlength{\parskip}{#2\baselineskip}
  \fi
 }%
 {}
\usepackage{calc} % for \widthof, \maxof
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\maxof{\widthof{#1}}{\csllabelwidth}}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth}{#1}}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}\geometry{verbose,letterpaper,tmargin=2.2cm,bmargin=2.2cm,lmargin=2.2cm,rmargin=2.2cm}

\usepackage{lineno}
\usepackage[nolists,noheads]{endfloat}

\pagestyle{plain}

\tolerance=1
\emergencystretch=\maxdimen
\hyphenpenalty=10000
\hbadness=10000

\doublespacing

\fancypagestyle{normal}
{
  \fancyhf{}
  \fancyfoot[R]{\footnotesize\sffamily\thepage\ of \pageref*{LastPage}}
}
\begin{document}
\raggedright
\thispagestyle{empty}
{\Large\bfseries\sffamily The missing link: discerning true from false
negatives when sampling species interaction networks}
\vskip 5em

%
\href{https://orcid.org/0000-0002-6506-6487}{Michael D.\,Catchen}%
%
\,\textsuperscript{1,2}\quad %
\href{https://orcid.org/0000-0002-0735-5184}{Timothée\,Poisot}%
%
\,\textsuperscript{3,2}\quad %
\href{https://orcid.org/0000-0002-6004-4027}{Laura\,Pollock}%
%
\,\textsuperscript{1,2}\quad %
\href{https://orcid.org/0000-0001-6075-8081}{Andrew\,Gonzalez}%
%
\,\textsuperscript{1,2}

\textsuperscript{1}\,McGill University\quad \textsuperscript{2}\,Québec
Centre for Biodiversity Sciences\quad \textsuperscript{3}\,Université de
Montréal

\textbf{Correspondance to:}\\
Michael D. Catchen --- \texttt{michael.catchen@mcgill.ca}\\

\vfill
This work is released by its authors under a CC-BY 4.0 license\hfill\ccby\\
Last revision: \emph{\today}

\clearpage
\thispagestyle{empty}

\vfill

        {\bfseries Abstract:}\,Ecosystems are composed of networks of
interacting species. These interactions allow communities of species to
persist through time through both neutral and adaptive processes. Still
a robust understanding of, and ability to predict and forecast
interactions among species remains elusive. This knowledge-gap is
largely driven by a shortfall of data---although species occurrence data
has rapidly increased in the last decade, species interaction data has
not kept pace, largely due to the intrinsic difficulty and effort
required to sample interactions. These sampling challenges bias data and
hinder inferences about the structure and dynamics of interactions
networks. Here, we demonstrate the realized false-negative rate (the
percentage of species that actually interact but for which we do not yet
have a record) can be quite high, even in thoroughly sampled systems,
due to the intrinsic variation in abundances across species in a
community. We illustrate how a null model of occurrence detection can be
used to estimate the false-negative rate in a given dataset. One
hypothesis is that interactions between ``rare'' species are themselves
rare because these species are less likely to encounter one-another than
species of higher relative abundance. However, we demonstrate that
across several datasets of spatial or temporally replicated networks,
there are positive associations that suggest these interactions actually
exist but just are not observed. Finally, we assess how false negatives
influence various models of network prediction, and recommend directly
accounting for observation error in predictive models. We conclude by
discussing how the understanding of false-negatives can inform how we
design monitoring schemes for species interactions.\\%

\vfill

\clearpage
\linenumbers
\pagestyle{normal}

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

Species interactions drive many processes in evolutionary biology and
community ecology. A better understanding of interactions among species
is an imperative to both mitigate the potentially harmful impacts of
anthropogenic change on Earth's biodiversity (Makiola \emph{et al.}
2020) and to predict zoonotic spillover of disease to prevent future
pandemics (Becker \emph{et al.} 2021). However, meeting these challenges
is difficult because interactions are intrinsically hard to sample
(Jordano 2016). Over the past few decades biodiversity data has become
increasingly available---for example, remote-sensing has enabled
collection of data on spatial scales and resolutions previously
unimaginable (Stephenson 2020), while the adoption of open data
practices (Kenall \emph{et al.} 2014) have substantially increased the
amount of data available to ecologists. Still, widespread data about
species interactions remains elusive (Poisot \emph{et al.} 2021).
Observing an interaction between two species often requires human
observation because remote sampling methods can primarily detect
co-occurrence (but see Niedballa \emph{et al.} (2019)), and this itself
is not necessarily indicative of an interaction (Blanchet \emph{et al.}
2020). This constraint induces biases on species interaction data
subject to the spatial and temporal scales that current observation
methods can feasibly sample. This is further compounded by semantic
confusion around the word ``interaction''---for example one might
consider competition a type of species interaction, even though it is
marked by a lack of co-occurrence between species, unlike other types of
interactions, like trophism or pollination, which require both species
to be together at the same place and time. We define interaction in the
latter sense, where two species have fitness consequences on one-another
if they are in the sample place at the same time. In addition, here we
only consider direct (not higher-order) interactions.

The importance of \emph{sampling effort} and its impact on resulting
ecological data has produced a rich body of literature. The recorded
number of species in a dataset or sample depends on the total number of
observations (Walther \emph{et al.} 1995; Willott 2001)---as do
estimates of population abundance (Griffiths 1998)---in addition to
spatial coverage and species detectability. This has motivated more
quantitatively robust approaches to account for error in sampling data
in many contexts: to determine if a given species is extinct (Boakes
\emph{et al.} 2015), to determine sampling design (Moore \& McCarthy
2016), and to measure species richness across large scales (Carlson
\emph{et al.} 2020). In the context of interactions, an initial concern
was the compounding effects of limited sampling effort combined with the
amalgamation of data (across both study sites, time of year, and
taxonomic scales) could lead any empirical set of observations to
inadequately reflect the reality of how species interact (Paine 1988) or
the structure of the network as a whole (Martinez \emph{et al.} 1999;
McLeod \emph{et al.} 2021). Martinez \emph{et al.} (1999) showed that in
a plant-endophyte trophic network, network connectance is robust to
sampling effort, but this was done in the context of a system for which
observation of 62,000 total interactions derived from 164,000
plant-stems was feasible. In some systems (e.g.~megafauna food-webs)
this many observations is either impractical or infeasible due to the
absolute abundance of the species in question.

We cannot feasibly observe all (or even most) of the interactions that
occur in an ecosystem. This means we can be confident two species
actually interact if we have a record of it (given an estimate of
species misidentification probability), but not at all confident that a
pair of species \emph{do not} interact if we have \emph{no record} of
those species observed together. In other words, it is difficult to
distinguish \emph{true-negatives} (two species never interact) from
\emph{false-negatives} (two species interact sometimes, but we do not
have a record of it). For a concrete example of a false-negative in a
food web, see fig.~\ref{fig:concept}. Because even the most highly
sampled systems will still contain missing interactions, there is
increasing interest in combining species-level data (e.g.~traits,
abundance, range, phylogenetic relatedness, etc.) to build models to
predict interactions between species we haven't observed together before
(Strydom \emph{et al.} 2021). However, the noise of false-negatives
could impact the efficacy of our predictive models and have practical
consequences for answering questions about interactions (de Aguiar
\emph{et al.} 2019). This data constraint is amplified as the
interaction data we have is geographically biased toward the usual
suspects (Poisot \emph{et al.} 2021). We therefore need a statistical
approach to assessing these biases in the observation process and their
consequences for our understanding of interaction networks.

The intrinsic properties of ecological communities create several
challenges for sampling: first, species are not observed with equal
probability---we are much more likely to observe a species of high
abundance than one of very low abundance (Poisot \emph{et al.} 2015).
Canard \emph{et al.} (2012) presents a null model of food-web structure
where species encounter one-another in proportion to each species'
relative-abundance. This assumes that there are no associations in
species co-occurrence due to an interaction (perhaps because this
interaction is ``important'' for both species; Cazelles \emph{et al.}
(2016)), but in this paper we later show increasing strength of
associations leads to increasing probability of false-negatives in
interaction data, and that these positive associations are rampant in
existing network data. Second, observed co-occurrence is often equated
with meaningful interaction strength, but this is not necessarily the
case (Blanchet \emph{et al.} 2020)---a true ``non-interaction'' would
require that neither of two species, regardless of whether they
co-occur, ever exhibit any meaningful effect on the fitness of the
other. So, although co-occurrence is not directly indicative of an
interaction, it \emph{is} a precondition for an interaction.

Here, we illustrate how our confidence that a pair of species never
interacts highly depends on sampling effort, and suggest that surveys of
species interactions can benefit from simulation modeling of the
sampling process. We demonstrate that the realized false-negative rate
of interactions is directly related to the relative abundance of the
species pool, and demonstrate how simulation can be used to produce a
null estimate of the false-negative rate given total sampling effort
(the total count of all individuals of all species seen). We show that
positive associations in co-occurrence data can increase the realized
number of false-negatives, and demonstrate these positive associations
are rampant in network datasets. We conclude by recommending that the
simulation of sampling effort and species occurrence can and should be
used to help design surveys of species interaction diversity (Moore \&
McCarthy 2016), and by advocating use of null models like those
presented here as a tool for both guiding design of surveys of species
interactions and for modeling detection error in predictive models.

\begin{figure}
\hypertarget{fig:concept}{%
\centering
\includegraphics{./figures/concept.png}
\caption{This conceptual example considers a sample of the trophic
community of bears, wolves, salmon (pink fish), pike (yellow fish),
berry trees, and aspen trees. The true metaweb (all realized
interactions across the entire spatial extent) is shown on the left. In
the center is what a hypothetical ecologist samples at each site. Notice
that although bears are observed co-occurring with both salmon and pike,
there was never a direct observation of bears eating pike, even though
they actually do. Therefore, this interaction between bears and pike is
a false negative.}\label{fig:concept}
}
\end{figure}

\hypertarget{how-many-observations-of-a-non-interaction-do-we-need-to-be-confident-its-a-true-negative}{%
\section{How many observations of a non-interaction do we need to be
confident it's a true
negative?}\label{how-many-observations-of-a-non-interaction-do-we-need-to-be-confident-its-a-true-negative}}

We start with a naive model of interaction detection: we assume that
every interacting pair of species is incorrectly observed as
not-interacting with an independent and fixed probability, which we
denote \(p_fn\) and subsequently refer to as the False-Negative Rate
(FNR). If we observe the same species not-interacting \(N\) times, then
the probability of a true-negative (denoted \(p_{tn}\)) is given by
\(p_{tn}=1-(p_{fn})^N\). This relation (the probability-mass-function of
geometric distribution, a special case of the negative-binomial
distribution) is shown in fig.~\ref{fig:geometric}(A) for varying values
of \(p_{fn}\) and illustrates a fundamental link between our ability to
reliably say an interaction doesn't exist---\(p_{tn}\)---and the number
of times \(N\) we have observed a given species. In addition, note that
there is no non-zero \(p_{fn}\) for which we can ever \emph{prove} that
an interaction does not exist---no matter how many observations of
non-interactions \(N\) we have, \(p_{tn}<1\).

From fig.~\ref{fig:geometric}(A) it is clear that the more often we see
two species co-occurring, but \emph{not interacting}, the more likely
the interaction is a true-negative. This has several practical
consequences: first it means negatives taken outside the overlap of the
range of each species aren't informative because co-occurrence was not
possible, and therefore neither was an interaction. Second, we can use
this relation to compute the expected number of total observations
needed to obtain a ``goal'' number of observations of a particular pair
of species (fig.~\ref{fig:geometric}(B)). As an example, if we
hypothesize that \(A\) and \(B\) do not interact, and we want to see
species \(A\) and \(B\) both co-occurring and \emph{not interacting} 10
times to be confident this is a true negative, then we need an expected
1000 observations of all species if the relative abundances of \(A\) and
\(B\) are both \(0.1\).

Because the true FNR is latent, we can never actually be sure what the
actual number of false negatives in our data---however, we can use
simulation to estimate it for datasets of a given size using neutral
models of observation. If some of the ``worst-case'' FNRs presented in
fig.~\ref{fig:geometric}(A) seem unrealistically high, consider that
species are observed in proportion to their relative abundance. In the
next section we demonstrate that the distribution of abundance in
ecosystems can lead to very high realized values of FNR (\(p_{fn}\))
simply as an artifact of sampling effort.

\begin{figure}
\hypertarget{fig:geometric}{%
\centering
\includegraphics{./figures/fig1.png}
\caption{\textbf{(A)} The probability that an observed interaction is a
true negative (y-axis) given how many times it has been sampled as a
non-interaction (x-axis). Each color reflects a different value of
\(p_{fn}\), the false-negative rate (FNR)---this is effectively the cdf
of the geometric distribution. \textbf{(B)} The expected number of total
observations needed (colors) to observe 10 co-occurrences between a
species with relative abundance \(P(A)\) (x-axis) and a second species
with relative abundance \(P(Y)\). \textbf{(C)}: False negative rate
(y-axis) as a function of total sampling effort (x-axis) and network
size, computed using the method described above. For 500 independent
draws from the niche model (Williams \& Martinez (2000)) at varying
levels of species richness (colors) with connectance drawn according to
the flexible-links model (MacDonald \emph{et al.} (2020)) as described
in the main text. For each draw from the niche model, 200 sets of 1500
observations are simulated, for which each the mean false negative rate
at each observation-step is computed. Means denoted with points, with 1
in the first shade and 2 in the second. \textbf{(D)}: Same as
\textbf{(C)}, except using empirical food webs from Mangal database,
where richness. The outlier on \textbf{(D)} is a 714 species
food-web.}\label{fig:geometric}
}
\end{figure}

\hypertarget{false-negatives-as-a-product-of-relative-abundance}{%
\section{False-negatives as a product of relative
abundance}\label{false-negatives-as-a-product-of-relative-abundance}}

We now show that the realized FNR changes drastically with sampling
effort due to the intrinsic variation of the abundance of individuals of
each species within a community. We do this by simulating the process of
observation of species interactions, applied both to 243 empirical food
webs from the Mangal database (Banville \emph{et al.} 2021) and random
food-webs generated using the niche model, a simple generative model of
food-web structure that accounts for allometric scaling (Williams \&
Martinez 2000). Our neutral model of observation assumes each observed
species is drawn in proportion to each species' abundance at that place
and time. The abundance distribution of a community can be
reasonably-well described by a log-normal distribution (Volkov \emph{et
al.} 2003). In addition to the log-normal distribution, we also tested
the case where the abundance distribution is derived from power-law
scaling \(Z^{(log(T_i)-1)}\) where \(T_i\) is the trophic level of
species \(i\) and \(Z\) is a scaling coefficient (Savage \emph{et al.}
2004), which yields the same qualitative behavior. The practical
consequence of abundance distributions spanning many orders of magnitude
of abundance is that observing two ``rare'' species interacting requires
two low probability events: observing two rare species \emph{at the same
time}.

To simulate the process of observation, for an ecological network \(M\)
with \(S\) species, we sample abundances for each species from a
standard-log-normal distribution. For each true interaction in the
adjacency matrix \(M\) (i.e. \(M_{ij}=1\)) we estimate the probability
of observing both species \(i\) and \(j\) at a given place and time by
simulating \(n\) observations of all individuals of any a species, where
the species of the individual observed at the \(\{1,2,\dots,n\}\)-th
observation is drawn from the generated log-normal distribution of
abundances. For each pair of species \((i,j)\), if both \(i\) and \(j\)
are observed within the n-observations, the interaction is tallied as a
true positive if \(M_{ij}=1\). If only one of \(i\) or \(j\) are
observed---but not both---in these \(n\) observations, but \(M_{ij}=1\),
this is counted as a false-negative, and a true-negative otherwise. For
each pair of species \((i,j)\), if both \(i\) and \(j\) are observed
within the n-observations, the interaction is tallied as a true positive
if \(M_{ij}=1\). If only one of \(i\) or \(j\) are observed---but not
both---in these \(n\) observations, but \(M_{ij}=1\), this is counted as
a false-negative, and a true-negative otherwise (\(M_{ij} = 0\)).

In fig.~\ref{fig:geometric}(C) we see this model of observation applied
to niche model networks across varying levels of species richness, and
in fig.~\ref{fig:geometric}(D) the observation model applied to Mangal
food webs. For all niche model simulations in this manuscript, for a
given number of species \(S\) the number of interactions is drawn from
the flexible-links model fit to Mangal data (MacDonald \emph{et al.}
2020), effectively drawing the number of interactions \(L\) for a random
niche model food-web as

\[L \sim  \text{BetaBinomial}(S^2-S+1,\mu\phi, 1-\mu\phi)\]

where the MAP estimate of \((\mu, \phi)\) applied to Mangal data from
(MacDonald \emph{et al.} 2020) is \((\mu=0.086, \phi=24.3)\). All
simulations were done with 500 independent replicates of unique niche
model networks per unique number of observations \(n\). All analyses
presented here are done in Julia v1.8 (Bezanson \emph{et al.} 2015)
using both EcologicalNetworks.jl v0.5 and Mangal.jl v0.4 (Banville
\emph{et al.} 2021) and are hosted at (GITHUB\_LINK\_TODO). Note that
the empirical data, for the reasons described above, very likely already
contains many false negatives, we'll revisit this issue in the final
section.

From fig.~\ref{fig:geometric}(C) it is evident that the number of
species considered in a study is inseparable from the false-negative
rate in that study, and this effect should be taken into account when
designing samples of ecological networks in the future. We see a similar
qualitative pattern in fig.~\ref{fig:geometric}(D) where the FNR drops
off quickly as a function of observation effort, mediated by total
richness. The practical consequence of the bottom row of
fig.~\ref{fig:geometric} is whether the total number of observations of
all species (the x-axis) for the threshold FNR we deem acceptable (the
y-axis) is feasible. This raises two points: first, empirical data on
interactions are subject to the practical limitations of funding and
human-work hours, and therefore existing data tend to fall on the order
of hundreds or thousands observations of individuals per site. Clear
aggregation of data on sampling effort has proven difficult to find and
a meta-analysis of network data and sampling effort seems both pertinent
and necessary, in addition to the effects of aggregation of interactions
across taxonomic scales (Gauzens \emph{et al.} 2013; Giacomuzzo \&
Jordán 2021). This inherent limitation on in-situ sampling means we
should optimize where we sample across space so that for a given number
of samples, we obtain the maximum information possible. Second, what is
meant by ``acceptable'' FNR? This raises the question: does a shifting
FNR lead to rapid transitions in our ability inference and predictions
about the structure and dynamics of networks, or does it produce a
roughly linear decay in model efficacy? We explore this in the next
section.

We conclude this section by advocating for the use of neutral models
similar to above to generate expectations about the number of
false-negatives in a data set of a given size. This could prove fruitful
both for designing surveys of interactions but also because we may want
to incorporate models of imperfect detection error into predictive
interactions models, as Joseph (2020) does for species occurrence
modeling. Additionally, we emphasize that one must consider the context
for sampling---is the goal to detect a particular species (as in
fig.~\ref{fig:geometric}(C)), or to get a representative sample of
interactions across the species pool? These arguments are
well-considered when sampling individual species (Willott 2001), but
have not yet been adopted for designing samples of communities.

\hypertarget{positive-associations-increase-the-false-negative-rate}{%
\section{Positive associations increase the false-negative
rate}\label{positive-associations-increase-the-false-negative-rate}}

This model above doesn't consider the possibility that there are
positive or negative associations which shift the probability of
observing two species together due to their interaction (Cazelles
\emph{et al.} 2016). However, here we demonstrate that the probability
of observing a false negative can be higher if there is some positive
association in the occurrence of species \(A\) and \(B\). If we denote
the probability that we observe the co-occurrence of two species \(A\)
and \(B\) as \(P(AB)\) and if there is no association between the
marginal probabilities of observing \(A\) and observing \(B\), denoted
\(P(A)\) and \(P(B)\) respectively, then the probability of observing
their co-occurrence is the product of the marginal probabilities for
each species, \(P(AB) = P(A)P(B)\). In the other case where there is
some positive strength of association between observing both \(A\) and
\(B\) because this interaction is ``important'' for each species, then
the probability of observation both \(A\) and \(B\), \(P(AB)\), is
greater than \(P(A)P(B)\) as \(P(A)\) and \(P(B)\) are not independent
and instead are positively correlated, i.e.~\(P(AB)> P(A)P(B)\). In this
case, the probability of observing a single false-negative in our naive
model from fig.~\ref{fig:geometric}(A) is \(p_{fn}= 1-P(AB)\), which due
to the above inequality implies \(p_{fn}>1-P(A)P(B)\). This indicates an
increasingly greater probability of a false negative as the strength of
association gets stronger, \(P(AB) \to P(AB) \gg P(A)P(B)\). However,
this still does not consider variation in species abundance in space and
time (Poisot \emph{et al.} 2015). If positive or negative associations
between species structure variation in the distribution of \(P(AB)\)
across space/time, then the spatial/temporal biases induced by data
collection would further impact the realized false negative rate, as the
probability of false negative would not be constant for each pair of
species across sites.

To test for these positive associations in data we scoured Mangal for
datasets with many spatial or temporal replicates of the same system.
For each dataset, we compute the marginal probability \(P(A)\) of
occurrence of each species \(A\) across all networks in the dataset. For
each pair of interacting species \(A\) and \(B\), we then compute and
compare the probability of co-occurrence if each species occurs
independently, \(P(A)P(B)\), to the empirical joint probability of
co-occurrence, \(P(AB)\). Following our analysis above, if \(P(AB)\) is
greater than \(P(A)P(B)\), then we expect our neutral estimates of the
FNR above to underestimate the realized FNR. In fig.~\ref{fig:mangal},
we see the difference between \(P(AB)\) and \(P(A)P(B)\) for the seven
suitable datasets with enough spatio-temporal replicates and a shared
taxonomic backbone (meaning all individual networks use common species
identifiers) found on Mangal to perform this analysis. Further details
about each dataset are reported in tbl.~\ref{tbl:id}.

In each of these datasets, the joint probability of co-occurrence
\(P(AB)\) is decisively greater than our expectation if species co-occur
in proportion to their relative abundance \(P(A)P(B)\). This suggests
that there may not be as many ``neutrally forbidden links'' (Canard
\emph{et al.} 2012) as we might think, and that the reason we do not
have records of interactions between rare species is probably due to
observation error. This has serious ramifications for the widely
observed property of nestedness seen in bipartite networks (Bascompte \&
Jordano 2007)---perhaps the reason we have lots of observations between
generalists is because they are more abundant, and this is particularly
relevant as we have strong evidence that generalism drives abundance
(Song \emph{et al.} 2022), not vice-versa.

\begin{figure}
\hypertarget{fig:mangal}{%
\centering
\includegraphics{./figures/fig2.png}
\caption{The difference between joint-probability of co-occurrence
(\(P(AB)\)) and expected probability of co-occurrence under independence
(\(P(A)P(B)\)) for interacting species for each dataset. The red-dashed
line indicates 0 (no association). Each histogram represents a density,
meaning the area of the entire curve sums to 1. The continuous density
estimate (computed using local smoothing) is shown in grey. The p-value
on each plot is the result of a one-sided t-test comparing the mean of
each distribution to 0.}\label{fig:mangal}
}
\end{figure}

\hypertarget{tbl:id}{}
\begin{longtable}[]{@{}ccccccccc@{}}
\caption{\label{tbl:id}This table describes the datasets used in the
above analysis (Fig 2). The table reports the type of each dataset, the
total number of networks in each dataset \((N)\), the total species
richness in each dataset \((S)\), the connectance of each metaweb (all
interactions across the entire spatial-temporal extent) \((C)\), the
mean species richness across each local network \(\bar{S}\), the mean
connectance of each local network \(\bar{C}\), the mean
\(\beta\)-diversity among overlapping species across all pairs of
network species (\(\bar{\beta}_{OS}\)), and the mean \(\beta\)-diversity
among all species in the metaweb (\(\bar{\beta}_{WN}\)). Both metrics
are computed using KGL \(\beta\)-diversity (Koleff \emph{et al.}
2003)}\tabularnewline
\toprule
Network & Type & \(N\) & \(S\) & \(C\) & \(\bar{S}\) & \(\bar{C}\) &
\(\bar{\beta}_{OS}\) & \(\bar{\beta}_{WN}\)\tabularnewline
\midrule
\endfirsthead
\toprule
Network & Type & \(N\) & \(S\) & \(C\) & \(\bar{S}\) & \(\bar{C}\) &
\(\bar{\beta}_{OS}\) & \(\bar{\beta}_{WN}\)\tabularnewline
\midrule
\endhead
Kopelke \emph{et al.} (2017) & Food Web & 100 & 98 & 0.037 & 7.87 &
0.142 & 1.383 & 1.972\tabularnewline
Thompson \& Townsend (2000) & Food Web & 18 & 566 & 0.014 & 80.67 &
0.049 & 1.617 & 1.594\tabularnewline
Havens (1992) & Food Web & 50 & 188 & 0.065 & 33.58 & 0.099 & 1.468 &
1.881\tabularnewline
Ponisio \emph{et al.} (2017) & Pollinator & 100 & 226 & 0.079 & 23.0 &
0.056 & 1.436 & 1.870\tabularnewline
Hadfield \emph{et al.} (2014) & Host-Parasite & 51 & 327 & 0.085 & 32.71
& 0.337 & 1.477 & 1.952\tabularnewline
Closs \& Lake (1994) & Food Web & 12 & 61 & 0.14 & 29.09 & 0.080 & 1.736
& 1.864\tabularnewline
CaraDonna \emph{et al.} (2017) & Pollinator & 86 & 122 & 0.18 & 21.42 &
0.312 & 1.527 & 1.907\tabularnewline
\bottomrule
\end{longtable}

\hypertarget{the-impact-of-false-negatives-on-network-properties-and-prediction}{%
\section{The impact of false-negatives on network properties and
prediction}\label{the-impact-of-false-negatives-on-network-properties-and-prediction}}

Here, we assess the effect of false negatives on our ability to make
predictions about interactions, as well as their effect on network
structure. The prevalence of false-negatives in data is the catalyst for
interaction prediction in the first place, and as a result methods have
been proposed to counteract this bias (Stock \emph{et al.} 2017; Poisot
\emph{et al.} 2022). However, it is feasible that the FNR in a given
dataset is so high that it could induce too much noise for an
interaction prediction model to detect the signal of possible
interaction between species.

To test this we use the dataset from Hadfield \emph{et al.} (2014) that
describes host-parasite interaction networks sampled across 51 sites,
and the same method as Strydom \emph{et al.} (2021) to extract latent
features for each species in this dataset based on applying PCA to the
co-occurrence matrix. We then predict a metaweb (equivalent to
predicting true or false for an interaction between each species pair,
effectively a binary classification problem) from these species-level
features using four candidate models for binary classification---three
often used machine-learning (ML) methods (Boosted Regression Tree (BRT),
Random Forest (RF), Decision Tree (DT)), and one naive model from
classic statistics (Logistic Regression (LR)). Each of the ML models are
bootstrap aggregated (or bagged) with 100 replicates each. We partition
the data into 80-20 training-test split, and then seed the training data
with false negatives at varying rates, but crucially do nothing to the
test data. We fit all of these models using MLJ.jl, a high-level Julia
framework for a wide-variety of ML models (Blaom \emph{et al.} 2020). We
evaluate the efficacy of these models using two common measures of
binary classifier performance: the area under the receiver-operator
curve (ROC-AUC) and the area under the precision-recall curve (PR-AUC),
for more details see Poisot (2022). Here, PR-AUC is slightly more
relevant as it is a better indicator of prediction of false-negatives.
The results of these simulations are shown in
fig.~\ref{fig:addedfnr}(A\&B).

\begin{figure}
\hypertarget{fig:addedfnr}{%
\centering
\includegraphics{./figures/fig3.png}
\caption{\textbf{(A)} The area-under the receiver-operator curve
(ROC-AUC) and \textbf{(B)} The area-under the precision-recall curve
(PR-AUC; right) for each different predictive model (colors/shapes)
across a spectrum of the proportion of added false negatives (x-axis).
\textbf{(C)} The mean trophic-level of all species in a network
generated with the niche model across different species richnesses
(colors). For each value of the FNR, the mean trophic level was computed
across 50 replicates. The shaded region for each line is one
standard-deviation across those replicates.}\label{fig:addedfnr}
}
\end{figure}

One interesting result seen in fig.~\ref{fig:addedfnr}(A\&B) is that the
ROC-AUC value does not approach random in the same way the PR-AUC curve
does as we increase the added FNR. The reason for this is that ROC-AUC
is fundamentally not as useful a metric in assessing predictive capacity
as PR-AUC. As we keep adding more false-negatives, the network
eventually becomes a zeros matrix, and these models can still learn to
predict ``no-interaction'' for all possible species pairs, which does
far better than random guessing (ROC-AUC = 0.5) in terms the false
positive rate (one of the components of ROC-AUC). This highlights a more
broad issue of label class imbalance, meaning there are far more
non-interactions than interactions in data. A full treatment of the
importance of class-balance is outside the scope of this paper, but is
explored in-depth in Poisot (2022).

Although these ML models are surprisingly performant at link prediction
given their simplicity, there have been several major developments in
applying deep-learning methods to many tasks in network inference and
prediction---namely graph-representation learning (GRL, Khoshraftar \&
An (2022)) and graph convolutional networks (Zhang \emph{et al.} 2019).
At this time, these advances can not yet be applied to ecological
networks because they require far more data than we currently have. We
already have lots of features that could be used as inputs into these
models (i.e.~species level data about occurrence, genomes, abundance,
etc.), but our network datasets barely get into the hundreds of local
networks sampled across space and time (tbl.~\ref{tbl:id}). Once we
start to get into the thousands, these models will become more useful,
but this can only be done with systematic monitoring of interactions.
This again highlights the need to optimize our sampling effort to
maximize the amount of information contained in our data given the
expensive nature of sampling interactions.

We also consider how the FNR affects network properties. In
fig.~\ref{fig:addedfnr}(C) we see the mean trophic level across networks
simulated using the niche model (as above), across a spectrum of FNR
values. In addition to the clear dependence on richness, we see that
mean trophic level, despite varying widely between niche model
simulations, tends to be relatively robust to false negatives and does
not deviate widely from the true value until very large FNRs,
i.e.~\(p_{fn} > 0.7\). This is not entirely unsurprising. Removing links
randomly from a food-web is effectively the inverse problem of the
emergence of a giant component (more than half of the nodes are in a
connected network) in random graphs (see Li \emph{et al.} (2021) for a
thorough review). The primary difference being that we are removing
edges, not adding them, and thus we are witnessing the dissolution of a
giant component, rather than the emergence of one. Further applications
of percolation theory to the topology of ecological networks could
improve our understanding of how false-negatives impact the inferences
about the structure and dynamics on these networks.

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

Species interactions enable the persistence and functioning of
ecosystems, but our understanding of interactions is limited due to the
intrinsic difficulty of sampling. Here we have provided a null model for
the expected number of false-negatives in an interaction dataset. We
demonstrated that we expect many false-negatives in species interaction
datasets purely due to the intrinsic variation of abundances within a
community. We also, for the first time to our knowledge, measured the
strength of association between co-occurrence and interactions (Cazelles
\emph{et al.} 2016) across many empirical systems, and found that these
positive associations are both very common, and showed algebraically
that they increase the realized FNR. We have also shown that
false-negatives could further impact our ability to both predict
interactions and infer properties of the networks, which highlights the
need for further research into methods for correcting this bias in
existing data.

A better understanding of how false-negatives impact species interaction
data is a practical necessity---both for inference of network structure
and dynamics, but also for prediction of interactions by using species
level information. False-negatives could pose a problem for many forms
of inference in network ecology. For example, inferring the dynamic
stability of a network could be prone to error if the observed network
is not sampled ``enough.'' What exactly ``enough'' means is then
specific to the application, and should be assessed via methods like
those here when designing samples. Further, predictions about network
rewiring (Thompson \& Gonzalez 2017) due to range shifts in response to
climate change could be error-prone without accounting for interactions
that have not been observed but that still may become climatically
infeasible. As is evident from fig.~\ref{fig:geometric}(A), we can never
guarantee there are no false-negatives in data. In recent years, there
has been interest toward explicitly accounting for false-negatives in
models (Stock \emph{et al.} 2017; Young \emph{et al.} 2021), and a
predictive approach to networks---rather than expecting our samples to
fully capture all interactions (Strydom \emph{et al.} 2021). As a
result, better models for predicting interactions are needed for
interaction networks. This includes explicitly accounting for
observation error (Johnson \& Larremore 2021)---certain classes of
models have been used to reflect hidden states which account for
detection error in occupancy modeling (Joseph 2020), and could be
integrated in the predictive models of interactions in the future.

A brief caveat here is that we do not consider false-positives---in
large part false-positives can be explained by misidentification of
species, although this could be a relevant consideration in some cases.
The same logic that we apply to false-negatives could easily be applied
to false-positives, e.g.~that we can be much more confident that an
interaction is a true positive if we have observed it 50 times rather
than only once, and we could similarly model this using the geometric
distribution as in fig.~\ref{fig:geometric}(A). However, because
ecological networks are so sparse, there are far more negatives than
positives in the dataset, and therefore likely to be far more
false-negatives than false-positives in absolute terms.

This work has several practical consequences for the design of
interaction samples. Simulating the process of observation could be a
powerful tool for estimating the sampling effort required by a study
that takes relative abundance into account, and provides a null baseline
for expected FNR. It is necessary to take the size of the species pool
into account when deciding how many total samples is sufficient for an
``acceptable'' FNR (fig.~\ref{fig:geometric}(C \& D)). Further the
spatial and temporal turnover of interactions means any approach to
sampling prioritization must be spatiotemporal. We demonstrated earlier
that observed negatives outside of the range of both species aren't
informative, and therefore using species distribution models could aid
in this spatial prioritization of sampling sites.

Our work highlights the need for a quantitatively robust approach to
sampling design, both for interactions (Jordano 2016) and all other
aspects of biodiversity (Carlson \emph{et al.} 2020). As anthropogenic
forces create rapid shifts in our planet's climate and biosphere, this
is an imperative to maximize the amount of ecological information we get
in our finite samples, and make our inferences and decisions based on
this data as robust as possible. Where we choose to sample, and how
often we choose to sample there, has strong impacts on the inferences we
make from data. Incorporating a better understanding of sampling effort
and bias to the design of biodiversity monitoring systems, and the
inference and predictive models we apply to this data, is imperative in
understanding how biodiversity is changing, and making actionable
forecasts about the future of ecological interactions on our planet.

\hypertarget{references}{%
\section*{References}\label{references}}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\hypertarget{ref-Banville2021ManJl}{}%
Banville, F., Vissault, S. \& Poisot, T. (2021). Mangal.jl and
EcologicalNetworks.jl: Two complementary packages for analyzing
ecological networks in Julia. \emph{Journal of Open Source Software}, 6,
2721.

\leavevmode\hypertarget{ref-Bascompte2007PlaMut}{}%
Bascompte, J. \& Jordano, P. (2007). Plant-Animal Mutualistic Networks:
The Architecture of Biodiversity. \emph{Annual Review of Ecology,
Evolution, and Systematics}, 38, 567--593.

\leavevmode\hypertarget{ref-Becker2021OptPre}{}%
Becker, D.J., Albery, G.F., Sjodin, A.R., Poisot, T., Bergner, L.M.,
Dallas, T.A., \emph{et al.} (2021). Optimizing predictive models to
prioritize viral discovery in zoonotic reservoirs.

\leavevmode\hypertarget{ref-Bezanson2015JulFre}{}%
Bezanson, J., Edelman, A., Karpinski, S. \& Shah, V.B. (2015). Julia: A
Fresh Approach to Numerical Computing.

\leavevmode\hypertarget{ref-Blanchet2020CooNot}{}%
Blanchet, F.G., Cazelles, K. \& Gravel, D. (2020). Co-occurrence is not
evidence of ecological interactions. \emph{Ecology Letters}, 23,
1050--1063.

\leavevmode\hypertarget{ref-Blaom2020MljJul}{}%
Blaom, A.D., Kiraly, F., Lienart, T., Simillides, Y., Arenas, D. \&
Vollmer, S.J. (2020). MLJ: A Julia package for composable machine
learning. \emph{Journal of Open Source Software}, 5, 2704.

\leavevmode\hypertarget{ref-Boakes2015InfSpe}{}%
Boakes, E.H., Rout, T.M. \& Collen, B. (2015). Inferring species
extinction: The use of sighting records. \emph{Methods in Ecology and
Evolution}, 6, 678--687.

\leavevmode\hypertarget{ref-Canard2012EmeStr}{}%
Canard, E., Mouquet, N., Marescot, L., Gaston, K.J., Gravel, D. \&
Mouillot, D. (2012). Emergence of Structural Patterns in Neutral Trophic
Networks. \emph{PLOS ONE}, 7, e38295.

\leavevmode\hypertarget{ref-CaraDonna2017IntRew}{}%
CaraDonna, P.J., Petry, W.K., Brennan, R.M., Cunningham, J.L.,
Bronstein, J.L., Waser, N.M., \emph{et al.} (2017). Interaction rewiring
and the rapid turnover of plantpollinator networks. \emph{Ecology
Letters}, 20, 385--394.

\leavevmode\hypertarget{ref-Carlson2020WhaWou}{}%
Carlson, C.J., Dallas, T.A., Alexander, L.W., Phelan, A.L. \& Phillips,
A.J. (2020). What would it take to describe the global diversity of
parasites? \emph{Proceedings of the Royal Society B: Biological
Sciences}, 287, 20201841.

\leavevmode\hypertarget{ref-Cazelles2016TheSpe}{}%
Cazelles, K., Araújo, M.B., Mouquet, N. \& Gravel, D. (2016). A theory
for species co-occurrence in interaction networks. \emph{Theoretical
Ecology}, 9, 39--48.

\leavevmode\hypertarget{ref-Closs1994SpaTem}{}%
Closs, G.P. \& Lake, P.S. (1994). Spatial and Temporal Variation in the
Structure of an Intermittent-Stream Food Web. \emph{Ecological
Monographs}, 64, 1--21.

\leavevmode\hypertarget{ref-deAguiar2019RevBia}{}%
de Aguiar, M.A.M., Newman, E.A., Pires, M.M., Yeakel, J.D., Boettiger,
C., Burkle, L.A., \emph{et al.} (2019). Revealing biases in the sampling
of ecological interaction networks. \emph{PeerJ}, 7, e7566.

\leavevmode\hypertarget{ref-Gauzens2013FooAgg}{}%
Gauzens, B., Legendre, S., Lazzaro, X. \& Lacroix, G. (2013). Food-web
aggregation, methodological and functional issues. \emph{Oikos}, 122,
1606--1615.

\leavevmode\hypertarget{ref-Giacomuzzo2021FooWeb}{}%
Giacomuzzo, E. \& Jordán, F. (2021). Food web aggregation: Effects on
key positions. \emph{Oikos}, 130, 2170--2181.

\leavevmode\hypertarget{ref-Griffiths1998SamEff}{}%
Griffiths, D. (1998). Sampling effort, regression method, and the shape
and slope of sizeabundance relations. \emph{Journal of Animal Ecology},
67, 795--804.

\leavevmode\hypertarget{ref-Hadfield2014TalTwo}{}%
Hadfield, J.D., Krasnov, B.R., Poulin, R. \& Nakagawa, S. (2014). A Tale
of Two Phylogenies: Comparative Analyses of Ecological Interactions.
\emph{The American Naturalist}, 183, 174--187.

\leavevmode\hypertarget{ref-Havens1992ScaStr}{}%
Havens, K. (1992). Scale and Structure in Natural Food Webs.
\emph{Science}, 257, 1107--1109.

\leavevmode\hypertarget{ref-Johnson2021BayEst}{}%
Johnson, E.K. \& Larremore, D.B. (2021). Bayesian estimation of
population size and overlap from random subsamples.

\leavevmode\hypertarget{ref-Jordano2016SamNet}{}%
Jordano, P. (2016). Sampling networks of ecological interactions.
\emph{Functional Ecology}, 30, 1883--1893.

\leavevmode\hypertarget{ref-Joseph2020NeuHie}{}%
Joseph, M.B. (2020). Neural hierarchical models of ecological
populations. \emph{Ecology Letters}, 23, 734--747.

\leavevmode\hypertarget{ref-Kenall2014OpeFut}{}%
Kenall, A., Harold, S. \& Foote, C. (2014). An open future for
ecological and evolutionary data? \emph{BMC Evolutionary Biology}, 14,
66.

\leavevmode\hypertarget{ref-Khoshraftar2022SurGra}{}%
Khoshraftar, S. \& An, A. (2022). A Survey on Graph Representation
Learning Methods.

\leavevmode\hypertarget{ref-Koleff2003MeaBet}{}%
Koleff, P., Gaston, K.J. \& Lennon, J.J. (2003). Measuring beta
diversity for presenceabsence data. \emph{Journal of Animal Ecology},
72, 367--382.

\leavevmode\hypertarget{ref-Kopelke2017FooStr}{}%
Kopelke, J.-P., Nyman, T., Cazelles, K., Gravel, D., Vissault, S. \&
Roslin, T. (2017). Food-web structure of willow-galling sawflies and
their natural enemies across Europe. \emph{Ecology}, 98, 1730--1730.

\leavevmode\hypertarget{ref-Li2021PerCom}{}%
Li, M., Liu, R.-R., Lü, L., Hu, M.-B., Xu, S. \& Zhang, Y.-C. (2021).
Percolation on complex networks: Theory and application. \emph{Physics
Reports}, Percolation on complex networks: Theory and application, 907,
1--68.

\leavevmode\hypertarget{ref-MacDonald2020RevLin}{}%
MacDonald, A.A.M., Banville, F. \& Poisot, T. (2020). Revisiting the
Links-Species Scaling Relationship in Food Webs. \emph{Patterns}, 1.

\leavevmode\hypertarget{ref-Makiola2020KeyQue}{}%
Makiola, A., Compson, Z.G., Baird, D.J., Barnes, M.A., Boerlijst, S.P.,
Bouchez, A., \emph{et al.} (2020). Key Questions for Next-Generation
Biomonitoring. \emph{Frontiers in Environmental Science}, 7.

\leavevmode\hypertarget{ref-Martinez1999EffSam}{}%
Martinez, N.D., Hawkins, B.A., Dawah, H.A. \& Feifarek, B.P. (1999).
Effects of Sampling Effort on Characterization of Food-Web Structure.
\emph{Ecology}, 80, 1044--1055.

\leavevmode\hypertarget{ref-McLeod2021SamAsy}{}%
McLeod, A., Leroux, S.J., Gravel, D., Chu, C., Cirtwill, A.R., Fortin,
M.-J., \emph{et al.} (2021). Sampling and asymptotic network properties
of spatial multi-trophic networks. \emph{Oikos}, 130, 2250--2259.

\leavevmode\hypertarget{ref-Moore2016OptEco}{}%
Moore, A.L. \& McCarthy, M.A. (2016). Optimizing ecological survey
effort over space and time. \emph{Methods in Ecology and Evolution}, 7,
891--899.

\leavevmode\hypertarget{ref-Niedballa2019AssAna}{}%
Niedballa, J., Wilting, A., Sollmann, R., Hofer, H. \& Courtiol, A.
(2019). Assessing analytical methods for detecting spatiotemporal
interactions between species from camera trapping data. \emph{Remote
Sensing in Ecology and Conservation}, 5, 272--285.

\leavevmode\hypertarget{ref-Paine1988RoaMap}{}%
Paine, R.T. (1988). Road Maps of Interactions or Grist for Theoretical
Development? \emph{Ecology}, 69, 1648--1654.

\leavevmode\hypertarget{ref-Poisot2022GuiPre}{}%
Poisot, T. (2022). Guidelines for the prediction of species interactions
through binary classification.

\leavevmode\hypertarget{ref-Poisot2021GloKno}{}%
Poisot, T., Bergeron, G., Cazelles, K., Dallas, T., Gravel, D.,
MacDonald, A., \emph{et al.} (2021). Global knowledge gaps in species
interaction networks data. \emph{Journal of Biogeography}, 48,
1552--1563.

\leavevmode\hypertarget{ref-Poisot2022NetEmb}{}%
Poisot, T., Ouellet, M.-A., Mollentze, N., Farrell, M.J., Becker, D.J.,
Brierly, L., \emph{et al.} (2022). Network embedding unveils the hidden
interactions in the mammalian virome.

\leavevmode\hypertarget{ref-Poisot2015SpeWhy}{}%
Poisot, T., Stouffer, D.B. \& Gravel, D. (2015). Beyond species: Why
ecological interaction networks vary through space and time.
\emph{Oikos}, 124, 243--251.

\leavevmode\hypertarget{ref-Ponisio2017OppAtt}{}%
Ponisio, L.C., Gaiarsa, M.P. \& Kremen, C. (2017). Opportunistic
attachment assembles plantpollinator networks. \emph{Ecology Letters},
20, 1261--1272.

\leavevmode\hypertarget{ref-Savage2004EffBod}{}%
Savage, V.M., Gillooly, J.F., Brown, J.H., West, G.B. \& Charnov, E.L.
(2004). Effects of Body Size and Temperature on Population Growth.
\emph{The American Naturalist}, 163, 429--441.

\leavevmode\hypertarget{ref-Song2022GenDri}{}%
Song, C., Simmons, B.I., Fortin, M.-J. \& Gonzalez, A. (2022).
Generalism drives abundance: A computational causal discovery approach.
\emph{PLOS Computational Biology}, 18, e1010302.

\leavevmode\hypertarget{ref-Stephenson2020TecAdv}{}%
Stephenson, P. (2020). Technological advances in biodiversity
monitoring: Applicability, opportunities and challenges. \emph{Current
Opinion in Environmental Sustainability}, Open issue 2020 part A:
Technology Innovations and Environmental Sustainability in the
Anthropocene, 45, 36--41.

\leavevmode\hypertarget{ref-Stock2017LinFil}{}%
Stock, M., Poisot, T., Waegeman, W. \& De Baets, B. (2017). Linear
filtering reveals false negatives in species interaction data.
\emph{Scientific Reports}, 7, 45908.

\leavevmode\hypertarget{ref-Strydom2021RoaPre}{}%
Strydom, T., Catchen, M.D., Banville, F., Caron, D., Dansereau, G.,
Desjardins-Proulx, P., \emph{et al.} (2021). A roadmap towards
predicting species interaction networks (across space and time).
\emph{Philosophical Transactions of the Royal Society B: Biological
Sciences}, 376, 20210063.

\leavevmode\hypertarget{ref-Thompson2017DisGov}{}%
Thompson, P.L. \& Gonzalez, A. (2017). Dispersal governs the
reorganization of ecological networks under environmental change.
\emph{Nature Ecology \& Evolution}, 1, 1--8.

\leavevmode\hypertarget{ref-Thompson2000ResSol}{}%
Thompson, R.M. \& Townsend, C.R. (2000). Is resolution the solution?:
The effect of taxonomic resolution on the calculated properties of three
stream food webs. \emph{Freshwater Biology}, 44, 413--422.

\leavevmode\hypertarget{ref-Volkov2003NeuThe}{}%
Volkov, I., Banavar, J.R., Hubbell, S.P. \& Maritan, A. (2003). Neutral
theory and relative species abundance in ecology. \emph{Nature}, 424,
1035--1037.

\leavevmode\hypertarget{ref-Walther1995SamEff}{}%
Walther, B.A., Cotgreave, P., Price, R.D., Gregory, R.D. \& Clayton,
D.H. (1995). Sampling Effort and Parasite Species Richness.
\emph{Parasitology Today}, 11, 306--310.

\leavevmode\hypertarget{ref-Williams2000SimRul}{}%
Williams, R.J. \& Martinez, N.D. (2000). Simple rules yield complex food
webs. \emph{Nature}, 404, 180--183.

\leavevmode\hypertarget{ref-Willott2001SpeAcc}{}%
Willott, S.j. (2001). Species accumulation curves and the measure of
sampling effort. \emph{Journal of Applied Ecology}, 38, 484--486.

\leavevmode\hypertarget{ref-Young2021RecPla}{}%
Young, J.-G., Valdovinos, F.S. \& Newman, M.E.J. (2021). Reconstruction
of plantpollinator networks from observational data. \emph{Nature
Communications}, 12, 3911.

\leavevmode\hypertarget{ref-Zhang2019GraCon}{}%
Zhang, S., Tong, H., Xu, J. \& Maciejewski, R. (2019). Graph
convolutional networks: A comprehensive review. \emph{Computational
Social Networks}, 6, 11.

\end{CSLReferences}

\end{document}